On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}

### 2020-03-30 engineering.tex
first complete draft communication insecurity

While \emph{agents} and \emph{communication} between agents are described and
defined in Section~\ref{sec:theory}, the concepts of \emph{port} (depicted in
Figure~\ref{fig:ioports}), \emph{channel} (depicted in
Figure~\ref{fig:channel}), \emph{functional and physical architecture}, and
\emph{functional block} have not yet been defined with respect to the
$\abf$-theory.  Informally, a channel\footnote{for now, we rely on the
intuitive notion of channel as ``something'' that transfers information as
Assertions between output and input port, i.e. monodirectional} identifies the
place where the communication takes place.  In this work, we only consider
mono-directional channels and communication but the extension to bi-directional
channel can be considered as the union of two unidirectional channels. A
mono-directional channel is defined by the assertions sent or received (over
the channel).

@@

\Paragraph{Port Weaknesses}
From this definition, it follows that there exist only the following \emph{six} types of weaknesses, generating six types of insecure port in RCC5:
\begin{enumerate}[start=1, label={W\arabic*)}]
	\item \emph{drop port} ($s\dropport a$), where assertions reaches the port but do not pass the boundary of the agent (i.e. do not become belief of the agent)
	\item \emph{insertion port} ($s\insertport a$), where some new information is believed by $a$ as incoming from the port but $s$ didn't send it
	\item \emph{injection port} ($s\injectionport a$), where information coming from $s$ is substituted with new information which becomes believed by $a$
	\item \emph{selective port}, where some information passes the port and part is either:
	\begin{enumerate}[start=1, label={W4.\arabic*)}]
		\item dropped ($s\selectivedropport a$), 
		\item inserted ($s\selectiveinsertport a$), or
		\item injected ($s\selectiveinjectionport a$)
\end{enumerate}
\end{enumerate}

@@

\Paragraph{Communication Channels}
We start by considering the difference between a (communication)
mono-directional channel (channel from now on) and an agent, as we did for the
ports, since the $\abf$-theory is a theory of agents.  In fact, if a channel
were considered an agent (channel-agent) then the question would be how an
agent would transfer its Assertions to the channel-agent. If the channel
between the agent and the channel-agent is again an agent, we would generate an
\emph{infinite regress}. Therefore, we do allow channel-agents but we assume
a finite depth (of detail) for a channel, where there exists a bottom-channel 
which is not an agent. For now, we do not constrain a channel-agent in any way
so there is no difference between a channel agent and agent. Therefore, for
the sake of simplicity, we consider, in this text, channels to be bottom-channels,
defined as agents with the pre-defined behavior of storing Assertions that are transferred
over them (i.e. over the channel we are defining).

\begin{definition}{\bf Mono-directional Channel --}\label{def:monochannel}
	Given two agents $s$ (sender) and $r$ (receiver) a
	mono-directional channel $s \rightarrow r$ is defined as the union of all the assertions 
	$\rassert{s}{r}$ sent by $s$ to $r$ ($\assertionRegion_s$) and received by $r$ from $s$ ($\assertionRegion_r$) (for any given instance of time). 
\end{definition}
The \emph{quality of a mono-directional} channel is defined as the 
$\rcc$ relation between the Assertions made by the sender and the ones received by the receiver, i.e. $\Rcc{\assertionRegion_s}{\beliefRegion_r}$.

\Paragraph{Channel Weaknesses}

\Paragraph{Security Requirements} 
While standards (such
as IEC62443-1-3\fixnote{mr}{check the correct version}) defines requirements as
``information in transit/at-rest should be encrypted'' we believe that, at
specification level, security requirements can be formulated as subcategories of
the well-known CIA triad (where CIA stands for Confidentiality, Integrity,
Availability).
Infosec, or Information Security, defines the security risks related to
information with the de-facto standard CIA triad , which is often criticized
\autocite{CIAcriticismCPS} for being too general or non-adequate (e.g. by
adding authenticity which is often used as a building block for confidentiality
and integrity) to be effectively applied to the engineering of systems. Due to
this, many researchers an organization tried to improve the CIA triad.  The
evolutions/extensions of the CIA triad has been documented, e.g., in
\autocite{Samonas2014cia} and is summarized in Table~\ref{tab:ciaevolution}.

@@

An representative example is the effort made by the OECD (Organisation for Economic
Co-operation and Development) in \autocite{OECD2013guidelines} to define
security guidelines for information system (1992) based on new principles such
as ``awareness, ethics, risk assessment'' and maintain those revising the
document (e.g.  following the multistakeholder expert consultation in 2013).
Other similar efforts focused on defining entirely new principles or extending
the CIA triad, such as \autocite{NISTSP800-160}. What is of interest for our
argument is that, to the best of our knowledge, all of the approaches that aim
at improving the CIA triad are besed on motivations related to empirical evidences. On the contrary,
we now define (in Section~\ref{sec:properties}) the CIA triad for an agent
defined in the $\abftheory$ theory and we test if (i) other properties are
allowed in the $\abftheory$ theory, and (ii) if and how the CIA triad can be
detailed in the $\abftheory$ theory.

@@

We now want to generalize the CIA triad to agents and not just 
information and map it to the $\abftheory$ theory.
\begin{definition}{\bf Infosec-Confidentiality --}\label{def:infosec-c}
	An Information $\varphi$ is considered to be confidential to a set of 
	agents $Ag_c$ if, per each agent $a\in Ag_c$, the content of the
	information (i.e. the data it carries) can only be understood (i.e. the relation
	between the information and its data can be known) by the
	agents in $Ag_c$.

>>>>

While \emph{agents} and \emph{communication} between agents are described and
defined in Section~\ref{sec:theory}, the concepts of \emph{port} (depicted in
Figure~\ref{fig:ioports}), \emph{channel} (depicted in
Figure~\ref{fig:channel}), \emph{functional and physical architecture}, and
\emph{functional block} have not yet been defined with respect to the
$\abf$-theory.  Informally, a channel\footnote{for now, we rely on the
intuitive notion of channel as ``something'' that transfers information as
Assertions between output and input port, i.e. mono-directional} identifies the
place where the communication takes place.  In this work, we only consider
mono-directional channels and communication but the extension to bi-directional
channel can be considered as the union of two unidirectional channels. A
mono-directional channel is defined by the assertions sent or received (over
the channel).

@@

\Paragraph{Port Weaknesses}
From this definition, it follows that there exist only the following \emph{six} types of weaknesses, generating six types of insecure port in RCC5 (the notation is reported for input-ports, i.e. on the lef-hand side of the arrow):
\begin{enumerate}[start=1, label={W\arabic*)}]
	\item \emph{drop port} ($s\dropport a$), where assertions reaches the port but do not pass the boundary of the agent (i.e. do not become belief of the agent)
	\item \emph{insertion port} ($s\insertport a$), where some new information is believed by $a$ as incoming from the port but $s$ didn't send it
	\item \emph{injection port} ($s\injectionport a$), where information coming from $s$ is substituted with new information which becomes believed by $a$
	\item \emph{selective port}, where some information passes the port and part is either:
	\begin{enumerate}[start=1, label={W4.\arabic*)}]
		\item dropped ($s\selectivedropport a$), 
		\item inserted ($s\selectiveinsertport a$), or
		\item injected ($s\selectiveinjectionport a$)
\end{enumerate}
\end{enumerate}

@@

\Paragraph{Communication Channels}
We start by considering the difference between a (communication)
mono-directional channel (channel from now on) and an agent, as we did for the
ports, since the $\abf$-theory is a theory of agents.  In fact, if a channel
were considered an agent (channel-agent) then the question would be how an
agent would transfer its Assertions to the channel-agent. If the channel
between the agent and the channel-agent is again an agent, we would generate an
\emph{infinite regress}. Therefore, we do allow channel-agents but we assume a
finite depth (of detail) for a channel, where there exists a bottom-channel
which is not an agent. For now, we do not constrain a channel-agent in any way
so there is no difference between a channel agent and agent. Therefore, for the
sake of simplicity, we consider, in this text, channels to be bottom-channels,
defined as agents with the pre-defined behavior (i.e. defined in a dogmatic
way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
prevents us from introducing additional constraints to the channel as storing
Assertions that are transferred over the channel, or filter out some
input-assertions.}.

\begin{definition}{\bf Secure Mono-directional Channel (bottom-channel) --}\label{def:monochannel}
	A mono-directional channel between the two agents ($s \rightarrow r$) is defined as 
	an agent which behavior is (dogmatically) defined as: to forward any Assertion received from $s$ over an input-port, to  
	the output-port where $r$ is listening to.
\end{definition}
The \emph{quality of a mono-directional} channel is defined as the 
$\rcc$ relation between the Assertions made by the sender and the ones received by the receiver, i.e. $\Rcc{\assertionRegion_s}{\assertionRegion_r}$.

\Paragraph{Channel Weaknesses}
Given that a mono-directional bottom-channel is assumed to be perfectly
forwarding any Assertion from its input-port to its output-port, there is no
insecure behavior but only the combination of the weaknesses of the input and
output port; therefore there exists $(7^2)-2=47$ theoretical configurations
($7^2$ because there are $6$ insecure types of port, plus $1$ secure type, on both input and oupput
side; and we exclude the configuration with $2$ secure types as input
and ouput, $-2$); where only 43 are possible.

\begin{enumerate}[start=5, label={W\arabic*)}]
	\item \emph{secure input port} and \emph{output drop port} ($s\sdcom r$), 
	\item \emph{secure input port} and \emph{output insertion port} ($s\sicom r$), 
	\item \emph{secure input port} and \emph{output injection port} ($s\sjcom r$), 
	\item \emph{secure input port} and \emph{output selective drop port} ($s\ssdcom r$), 
	\item \emph{secure input port} and \emph{output selective insertion port} ($s\ssicom r$), 
	\item \emph{secure input port} and \emph{output selective injection port} ($s\ssjcom r$),

	\item \emph{input drop port} and \emph{output drop port} ($s\ddcom r$)
	\item \emph{input drop port} and \emph{output insertion port} ($s\dicom r$)
	\item \emph{input drop port} and \emph{output secure port} ($s\dscom r$)

	\item \emph{input insert port} and \emph{output drop port} ($s\idcom r$)
	\item \emph{input insert port} and \emph{output insert port} ($s\iicom r$)
	\item \emph{input insert port} and \emph{output injection port} ($s\ijcom r$)
	\item \emph{input insert port} and \emph{output selective drop port} ($s\isdcom r$)
	\item \emph{input insert port} and \emph{output selective insertion port} ($s\isicom r$)
	\item \emph{input insert port} and \emph{output selective injection port} ($s\isjcom r$)
	\item \emph{input insert port} and \emph{output secure port} ($s\iscom r$)

	\item \emph{input injection port} and \emph{output drop port} ($s\jdcom r$)
	\item \emph{input injection port} and \emph{output insertion port} ($s\jicom r$)
	\item \emph{input injection port} and \emph{output injection port} ($s\jjcom r$)
	\item \emph{input injection port} and \emph{output selective drop port} ($s\jsdcom r$)
	\item \emph{input injection port} and \emph{output selective insertion port} ($s\jsicom r$)
	\item \emph{input injection port} and \emph{output selective injection port} ($s\jsjcom r$)
	\item \emph{input injection port} and \emph{output secure port} ($s\jscom r$)

	\item \emph{input selective drop port} and \emph{output drop port} ($s\sddcom r$)
	\item \emph{input selective drop port} and \emph{output insertion port} ($s\sdicom r$)
	\item \emph{input selective drop port} and \emph{output injection port} ($s\sdjcom r$)
	\item \emph{input selective drop port} and \emph{output selective drop port} ($s\sdsdcom r$)
	\item \emph{input selective drop port} and \emph{output selective insertion port} ($s\sdsicom r$)
	\item \emph{input selective drop port} and \emph{output selective injection port} ($s\sdsjcom r$)
	\item \emph{input selective drop port} and \emph{output secure port} ($s\sdscom r$)

	\item \emph{input selective insertion port} and \emph{output drop port} ($s\sidcom r$)
	\item \emph{input selective insertion port} and \emph{output insertion port} ($s\siicom r$)
	\item \emph{input selective insertion port} and \emph{output injection port} ($s\sijcom r$)
	\item \emph{input selective insertion port} and \emph{output selective drop port} ($s\sisdcom r$)
	\item \emph{input selective insertion port} and \emph{output selective insertion port} ($s\sisicom r$)
	\item \emph{input selective insertion port} and \emph{output selective injection port} ($s\sisjcom r$)
	\item \emph{input selective insertion port} and \emph{output secure port} ($s\siscom r$)

	\item \emph{input selective injection port} and \emph{output drop port} ($s\sjdcom r$)
	\item \emph{input selective injection port} and \emph{output insertion port} ($s\sjicom r$)
	\item \emph{input selective injection port} and \emph{output injection port} ($s\sjjcom r$)
	\item \emph{input selective injection port} and \emph{output selective drop port} ($s\sjsdcom r$)
	\item \emph{input selective injection port} and \emph{output selective insertion port} ($s\sjsicom r$)
	\item \emph{input selective injection port} and \emph{output selective injection port} ($s\sjsjcom r$)
	\item \emph{input selective injection port} and \emph{output secure port}  ($s\sjcom r$)
\end{enumerate}
In Appendix~\ref{app:insecurecom}, we report the proof which is exhaustive on all
possible combinations. The only

\Paragraph{Security Requirements} 
While standards (such
as IEC62443-1-3\fixnote{mr}{check the correct version}) defines requirements as
``information in transit/at-rest should be encrypted'' we believe that, at
specification level, security requirements can be formulated as subcategories of
the well-known CIA triad (where CIA stands for Confidentiality, Integrity,
Availability).
Infosec, or Information Security, defines the security risks related to
information with the de-facto standard CIA triad , which is often criticized
\autocite{CIAcriticismCPS} for being too general or non-adequate (e.g. by
adding authenticity which is often used as a building block for confidentiality
and integrity) to be effectively applied to the engineering of systems. Due to
this, many researchers and organization tried to improve the CIA triad.  The
evolutions/extensions of the CIA triad has been documented, e.g., in
\autocite{Samonas2014cia} and is summarized in Table~\ref{tab:ciaevolution}.

@@

An representative example is the effort made by the OECD (Organisation for Economic
Co-operation and Development) in \autocite{OECD2013guidelines} to define
security guidelines for information system (1992) based on new principles such
as ``awareness, ethics, risk assessment'' and maintain those revising the
document (e.g.  following the multistakeholder expert consultation in 2013).
Other similar efforts focused on defining entirely new principles or extending
the CIA triad, such as \autocite{NISTSP800-160}. What is of interest for our
argument is that, to the best of our knowledge, all of the approaches that aim
at improving the CIA triad are besed on motivations related to empirical evidences. On the contrary,
we now define (in Section~\ref{sec:properties}) the CIA triad for an agent
defined in the $\abftheory$-theory and we test if (i) other properties are
allowed in the $\abftheory$-theory, and (ii) if and how the CIA triad can be
detailed in the $\abftheory$-theory.

@@

We now want to generalize the CIA triad to agents and not just 
information and map it to the $\abftheory$-theory. For the sake of simplicity,
we revert the order and start from availability.

\begin{figure}[t]
	\centering
	\includegraphics[width=\textwidth]{client-server.pdf}
	\caption{Client-Server paradigm in $\abf$-theory}
	\label{fig:client-server}
\end{figure}
Availability, again, is the denial of use of a resource. In the $\abf$-theory, 
in order to prevent a resource to be available, there
exists only (i.e. inclusively) the following mutually exclusive cases
that we illustrate using a client-server paradigm (depicted in Figure~\ref{fig:client-server}).
\begin{enumerate}
	\item\label{case:srv1} The resource on the server cannot be found, so that it is
		possible for the client to contact the server and the server
		sends a response back, but the output-port does not receive the
		necessary information. This may be due to a software error such
		that the functional architecture doesn't output a response, or
		a problem in the flow of information (i.e. the functional
		architecture doesn't receive the necessary information from
		other resources), or an hardware failure \&c. This is formalized
		as the relation between the beliefs produced by the functional
		architecture and the ones received by the output port, i.e.
		$\rcc(\behaviorRegion_s,\{\text{res}\})$ or 
		$\rcc(\beliefRegion_s,\{\text{res}\})$ 
	\item\label{case:srv2} The server hosting the resource is down, and then the server
		cannot output responses (i.e. insecure output port due to drop-related issue) or receive
		requests (insecure input port due to drop-related issue) or both.
	\item The client cannot properly compute request (simmetrically to Case~\ref{case:srv1})
	\item The client is down, symmetric to Case~\ref{case:srv2}
	\item The channel is down due to a drop related issue in one of the two channels 
\end{enumerate}

\begin{definition}{\bf Infosec-Availability --}\label{def:infosec-c}
\end{definition}

\begin{definition}{\bf Agent-Availability --}\label{def:agent-a}
\end{definition}

\begin{definition}{\bf System-Availability --}\label{def:system-a}
\end{definition}

\begin{definition}{\bf Infosec-Confidentiality --}\label{def:infosec-c}
	An Information $\varphi$ is considered to be confidential to a set of 
	agents $Ag_c$ if, per each agent $a\in Ag_c$, the content of the
	information (i.e. the data it carries) can only be understood (i.e. the relation
	between the information and its data can be known) by the
	agents in $Ag_c$.
